The Bolocam Galactic Plane Survey: Survey Description and Data Reduction

### 2009-03-30 methods/ms.tex
One more time, baby ...

@@

\usepackage{rotating}

@@

\newcommand{\bgpsarea}{150}
\newcommand{\bcamfwhm}{31.2\arcsec}
\newcommand{\ncores}{$10^4$}
\newcommand{\bgpsdepthlow}{20}
\newcommand{\bgpsdepthhigh}{50}

@@

@@

\newcommand{\penn}{1}
\newcommand{\casa}{2}
\newcommand{\cso}{3}
\newcommand{\utexas}{4}
\newcommand{\virginia}{5}
\newcommand{\ubc}

@@

\author{James Aguirre\altaffilmark{\penn},
        Adam Ginsburg\altaffilmark{\casa},
        Miranda Nordhaus\altaffilmark{\utexas},
	Meredith Drosback\altaffilmark{4},
        John Bally\altaffilmark{\casa}, 
	Cara Battersby\altaffilmark{\casa},
	Eric Todd Bradley\altaffilmark{5},
%        Richard Chamberlin\altaffilmark{\cso},
	Claudia Cyganowski\altaffilmark{6},
	Darren Dowell\altaffilmark{7}
	Neal J. Evans II\altaffilmark{\utexas},
        Jason Glenn\altaffilmark{\casa},
        Paul Harvey\altaffilmark{1,7},
        Erik Rosolowsky\altaffilmark{8},
%        Wayne Schlingman\altaffilmark{9},
%        Yancy Shirley\altaffilmark{9},
        Guy S. Stringfellow\altaffilmark{\casa},
        Josh Walawender\altaffilmark{10}, and 
        Jonathan Williams\altaffilmark{11}
}

%\author{James Aguirre\altaffilmark{\penn},
%  Adam Ginsburg\altaffilmark{\casa},
%  John Bally\altaffilmark{\casa},

%  Meredith Drosback\altaffilmark{\virginia},
%  Neal J. Evans\altaffilmark{\utexas},
%  Jason Glenn\altaffilmark{\casa},
%  Miranda Nordhaus\altaffilmark{\utexas},
%  Erik Rosolowsky\altaffilmark{\ubc},
%  Josh Walawender\altaffilmark{\virginia},
%  Jonathan Williams\altaffilmark{\virginia}
%}

\email{jaguirre@sas.upenn.edu}

\altaffiltext{\penn}{University of Pennsylvania, Philadelphia, PA 19104}
%Jansky Fellow, National Radio Astronomy Observatory} 
\altaffiltext{\casa}{CASA, University of Colorado, CB 389, Boulder, CO 80309}
\altaffiltext{\utexas}{Department of Astronomy, University of Texas, 
       1 University Station C1400, Austin, TX 78712}
\altaffiltext{4}{Caltech Submillimeter Observatory, Hilo, HI}
\altaffiltext{5}{Institute for Astronomy (IfA), University of Hawaii
640 N. Aohoku Pl., Hilo, HI 96720}

\begin{abstract}

@@

The advent of focal plane arrays containing hundreds of individual
bolometers sensitive to millimeter and sub-millimeter (sub-mm)
radiation has enabled large-scale surveys of mm-wavelength continuum
emission from Galactic interstellar dust.  The 144 element Bolocam
focal plane array, mounted at the Cassegrain focus of the 10 meter
diameter sub-mm telescope at the Caltech Sub-millimeter Observatory
(CSO) was use to survey over \bgpsarea\ square degrees of the northern
Galactic plane at a wavelength of 1.1 mm in the dust-continuum.  The
Bolocam Galactic Plane Survey (BGPS) presented here has detected about
\ncores\ cores in the surveyed fields, providing an unbiased sample of
cores suitable for multi-wavelength and high-resolution studies with
existing telescopes and future facilities.  The core catalog is
described in a companion paper \citep{rosolowsky09}.

@@

The great advantage of continuum observations at 1.1mm is that the
mass column density does not depend strongly on the dust temperature
$T_d$ (varying only linearly unless $T_d$ is less than about 10 K) and
is independent of the gas volume density.  Using the same dust opacity
as \citet{enoch06} and \citet{young06}, and assuming a gas-to-dust
mass ratio $X$ of 100, the mass sensitivity can be written as
\begin{equation} 
\label{eq:Mass}
M_{gas}\approx
0.22\left( e^{12.9/T_d}-1 \right) \left({S_\nu\over 15\; {\rm mJy}} \right)
D^{2}_{kpc} \msol 
\end{equation}
where $D_{kpc}$ is the distance to the source in kpc.  (At $S_\nu =
15$~mJy this corresponds to an RMS extinction sensitivity of $A_V\sim
1$ mag, with the same $T_d$ dependence).  It is thus possible with a
wide area, relatively shallow 1.1~mm survey to provide an unmatched,
uniform inventory of massive star-forming and starless condensations.

A number of (sub)millimeter surveys of the Galactic Plane are ongoing
or planned.  The BGPS is the first.

@@

%\Figure{mass_limits}{The limiting gas mass of cores detectable
%with at a flux limit of 15 mJy RMS. Detection thresholds at 3 (dashed)
%and 5 (solid) $\sigma$ are shown for a range of dust temperatures.
%Dust temperatures $<20$~K are most interesting for this survey, since
%these are the sources most likely to be pre- or proto-stellar
%condensations.}{fig:mass_limits}{1.0}
%
\Figure{coverage_moll} {The coverage of the BGPS.  The background
greyscale is IRAS 100 \mum.  
%The Bolocam coverage is shown with white
%solid lines in the top two panels and black in the bottom three.
%Major regions of Galactic star formation are labeled.  The coverage of
%{\em Spitzer}-GLIMPSE \citep{benjamin03} and the FCRAO $^{13}$CO GRS
%\citep{jackson06} are shown.
}{fig:coverage}{1.0}

@@

Bolocam\footnote{{\tt http://www.cso.caltech.edu/bolocam}} is the
facility 144-element bolometer array camera operating from the 10.4 m
dish of the Caltech Submillimeter Observatory (CSO) on the summit of
Mauna Kea.  We used the filter configuration with a bandcenter of
268~GHz (hereafter 1.1~mm) and fractional bandwidth $\Delta \nu/\nu =
0.17$.  The passband is constructed to exclude the $^{12}\mathrm{CO}(2
\to 1)$ emission line.  Figure \ref{fig:Passband} shows the passband.
We compute color corrections for sources of varying spectral index in
Appendix \ref{app:ColorCorrections}.

@@

\Figure{bolocam_bandpass}{The Bolocam 1.1 mm bandpass.  Bright
molecular emission lines are shown at their rest frequencies.  Note
that the Bolocam passband rejects $>90\%$ of the $^{12}$CO flux, leaving our
continuum measurements largely uncontaminated.  SO and CH3OH lines are probably
the dominant contributors to line flux in the passband.  \citet{nummelin1998}
found that 22\% of the  flux in one pointing towards Sgr B2 was from lines, and
\citet{yoshida2005} reported >40\% in Orion A from line emission}
{fig:Passband}{1.0}

@@

\Figure{scan_10deg}{The effect of field rotation on the coverage
obtained via Bolocam raster scans.  (could also use this figure to
show the distortion map)}{fig:Array}{1.0}

%\begin{sidewaysfigure}[h]
\begin{center}
  \begin{figure}[t]
%    \includegraphics[width=6in,height=9.5in]{fig1_pg1_bw_inv_bar_label} 
    \includegraphics[scale=0.9]{fig1_pg1_bw_inv_bar_label}
    \caption{$l=-10.5$ to $l=19.5$ Units are Jy/Beam.  The brightest
      sources, e.g. Sgr B2, Sgr A, and sources near $l=10$ and
      $l=13$, appear to be saturated, but this is only a display
      artifact.  The noise is more pronounced from $l=-7$ to $l=-2$
      because this region was observed less.}
    \label{fig:galplane_p1}
  \end{figure}
    %\end{sidewaysfigure}

    %\begin{sidewaysfigure}[h]
  \begin{figure} 
%    \includegraphics[width=6in,height=9.5in]{fig1_pg2_bw_inv_bar_label} 
    \includegraphics[scale=0.9]{fig1_pg2_bw_inv_bar_label} 
    \caption{$l=19.5$ to $l=49.5$.  Units are Jy/Beam .  G34.3+0.15, W
      51, W 43, W 49, and M 17 appear to be saturated, but this is
      only a display artifact}
    \label{fig:galplane_p2}
    %\end{sidewaysfigure}
  \end{figure}
\end{center}

%\begin{sidewaysfigure}[h]
%\begin{center}
%\includegraphics[scale=0.5]{bgps_montage}
%\end{center}
%\caption{ The coverage of the Bolocam Northern Galactic Plane Survey.
%as of September 2007.  The negative greyscale is the IRAS 100 \mum\
%survey.  Dashed lines indicate coverage between 15 and 100 mJy RMS;
%solid lines indicate areas covered more deeply than 30 mJy RMS.}
%\label{fig:coverage_sep06}
%\end{sidewaysfigure}

%\include{bgps_methods_flux_calibration}
\section{Flux Density and Surface Brightness Calibration}

The absolute flux calibration is obtained from observations of Mars,
Uranus and Neptune (the ``primary calibrators'').  The millimeter-wave
flux of these planets is known to $\sim 5\%$
\citep{orton86,griffin93}.  Calibrations at (sub)millimeter
wavelengths are strongly affected by atmospheric opacity corrections.
Further, the detector responsivity of Bolocam's bolometers is a
non-linear function of the mean atmospheric loading.

@@

A fit to the observed values for the primary calibrators was performed
for the each epoch separately.
%CHECK
The agreement between epochs was good, so a single combined
calibration was used for all data here.
%We have found that the calibration can differ
%between successive cooldowns of the instrument by $\sim10\%$ \Check\
%but is generally stable over the period of order a month which
%constitutes each of the observing epochs.  
The calibration curve is shown in Figure \ref{fig:Calibration}.  The
resulting error on the calibration curve fit is less than 
%CHECK
$5\%$ (statistical) over the observed range of $\tau$.  A tally of all
contributions to the flux density error is given in Table
\ref{tab:ErrorBudget}.

The above flux density calibration only accounts for the average
calibration of bolometer Volts to Janskys.  It is also necessary to
account for the variation of bolometer response across the focal
plane.  We do this by monitoring the response to the atmosphere
emission in all bolometers.  Since the atmosphere is common to all
detectors, any changes in response are due to the intrinsic properties
of the detectors.  The change in relative response with loading for a
typical detector is shown in Figure \ref{fig:Flatfield}.  The method
is discussed further in Section \ref{sec:Mapping}

The instrument beam size (PSF) is measured using the planets Uranus
and Neptune which are nearly point sources for Bolocam.
% For

%The semi-diameter of
%Mars over this period ranged from 2.3\arcsec\ to 2.9\arcsec, with
%Neptune $\sim1.1\arcsec$.  Relative to the \bcamfwhm~FWHM Bolocam
%beam, these may be regarded as point sources; the finite size of Mars
%results in a $< 1.5\%$ correction of the observed flux.
The beam area is that averaged over all detectors, and is ?$1.71\times10^{-6}$\TBD\ ($35.6$\farcs)
steradians.  This allows conversion of the maps, calibrated in Jy
beam$^{-1}$, into surface brightness (MJy sr$^{-1}$).  The uncertainty
in the beam size is ?$.7\%$\TBD\  (number quoted is the uncertainty in the 
mean beam size), leading to an uncertainty in the surface
brightness calibration of \TBD.

@@

Comparison of the accuracy of the Bolocam flux density calibration
with the SCUBA 850 \mum\ results is complicated by the unknown effect
of the dust spectral index.  However, there are several large-area
surveys with the MAMBO instrument at 1.2 mm with which we can compare.
For example, there is \citet{motte07}.

\Figure{fluxcal_errorbars}
%{calibration_plot_2004}
{Average calibration curve [V/Jy] versus the mean detector voltage, a
proxy for atmospheric loading.}{fig:Calibration}{1.0}

\Figure{rel_flux_cal_050303_ob3_to_050405_o66}{Scaling of the
relative calibration with atmospheric loading for bolocam
143. {\em To be replaced with new figure}}{fig:Flatfield}{1.0}

\Figure{relsens_cal}{Relative sensitivity calibration.  Black is the median over
all bolometers, red and green are individual bolometers. a. before
b. after}{fig:relsens_cal}{1.0}

@@

We constructed an absolute reference system for the BGPS by observing
bright quasars and blazars near the Galactic Plane.

The distribution of sources over the sky is shown in Figure \ref{fig:astrometry}

@@

@@

\Figure{plane_astrometry}{The distribution of absolute reference
sources (crosses) and in-plane millimeter sources (x's) in the
Galactic Plane.}{fig:astrometry}{1.0}

\Figure{pointing_model}{The pointing model correction (left) and
residuals (right) for Epoch V, from which the master reference images
were derived for subsequent alignment. {\em To be replaced with newer
figure?}}{fig:PointingModel}{1.0}

\Figure{pointing_model_residuals}{The residuals of the pointing
model.  Note the Gaussian distribution.}{fig:PointingModelResiduals}{1.0}

\Figure{pointing_model_skydist}{The distribution of the PCAL
sources across the local sky in Hawaii.  Notice the good sampling of
the entire region used for the Survey.}{fig:PCALSkydist}{1.0}

@@

We proceed iteratively as follows. 
\begin{equation}
\label{eq:FirstIterStep}
\tilde{a}^{(n)}(t) = 
\sum_i{d_i(t)-\tilde{s}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)}
\end{equation}
where $\tilde{s}^{(0)}_i(t)=0$.  This mean atmosphere model is then
fit to each bolometer to obtain the relative gains (flat field) of the
detectors
\begin{equation}
\tilde{a}^{(n)}_i(t) = \frac{d_i(t) \cdot a(t)}{d_i(t)^2} a(t) = r_i a(t)
\end{equation}
The instrument errors, which are typically smaller than the
atmosphere, are then estimated
\begin{equation}
\tilde{e}^{(n)}_i(t) = some models
\end{equation}
and both the atmosphere and instrument models are subtracted from the
raw time series
\begin{equation}
d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)
\approx s_i(t) + \epsi_i(t)
\end{equation}
This is the best estimate of signal plus irreducible noise at
iteration step $n$, and is made into a map
\begin{equation}
M[d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)] = \tilde{m}^{(n)}
\end{equation}
The current best map $\tilde{m}^{(n)}$ is then deconvolved to provide
a relatively low-noise, smooth map from which to generate a timestream
\begin{equation}
T[{\cal D}[\tilde{m}^{(n)}]] = \tilde{s}^{(n)}_i(t)
\end{equation}
We note that it is possible to produce an ``error map''
\begin{equation}
N = M[d_i(t) - \tilde{s}^{(n)}_i(t) - \tilde{a}^{(n)}_i(t) - 
\tilde{e}^{(n)}_i(t)]
\end{equation}
The process then begins again with Equation \ref{eq:FirstIterStep}.

@@

\Figure{flagger_withglitch}{An illustration of the flagging process using the
waterfall plot.  In this image, an anomaly which affects all
bolometers for a brief period of time is evident near the top of the
image.}{fig:Flag1}{1.0}

@@

After this is done, the remaining

>>>>

% Figure list
% 1) fig:Coverage
% 2) fig:BGPSMontage
% 3) fig:Bandpass
% 4) fig:Array
% 5) fig:CalibrationCurves
% 6) fig:Flatfield
% 7) fig:CygnusComparison
% 8) fig:PointingCalibrators

@@

\usepackage{rotating}
\usepackage{subfigure}

@@

\def\bgps{BGPS}
\newcommand{\bgpsarea}{150}
\newcommand{\bcamfwhm}{31.2\arcsec}
\newcommand{\ncores}{$10^4$}
\newcommand{\bgpsdepthlow}{20}
\newcommand{\bgpsdepthhigh}{50}

@@

\def\FigureTwo#1#2#3#4#5{
\begin{figure}[htb]
\epsscale{#5}
\plottwo{#1}{#2}
\caption{#3}
\label{#4}
\end{figure}
}

@@

\newcommand{\penn}{1}
\newcommand{\casa}{2}
\newcommand{\utexas}{3}
\newcommand{\virginia}{4}
\newcommand{\ucf}{5}
\newcommand{\wisc}{6}
\newcommand{\jpl}{7}
\newcommand{\ubc}{8}
\newcommand{\hawaiihilo}{9}
\newcommand{\hawaiimain}{10}
\newcommand{\arizona}{11}

@@

\author{James Aguirre\altaffilmark{\penn},
        Adam Ginsburg\altaffilmark{\casa},
        Miranda Nordhaus\altaffilmark{\utexas},
	Meredith Drosback\altaffilmark{\virginia},
        John Bally\altaffilmark{\casa}, 
	Cara Battersby\altaffilmark{\casa},
	Eric Todd Bradley\altaffilmark{\ucf},
%        Richard Chamberlin\altaffilmark{\cso},
	Claudia Cyganowski\altaffilmark{\wisc},
	Darren Dowell\altaffilmark{\jpl}
	Neal J. Evans II\altaffilmark{\utexas},
        Jason Glenn\altaffilmark{\casa},
        Paul Harvey\altaffilmark{\utexas,\casa},
        Erik Rosolowsky\altaffilmark{\ubc},
%        Wayne Schlingman\altaffilmark{9},
%        Yancy Shirley\altaffilmark{9},
        Guy S. Stringfellow\altaffilmark{\casa},
        Josh Walawender\altaffilmark{\hawaiihilo}, and 
        Jonathan Williams\altaffilmark{\hawaiimain}
}

\affil{{$^\penn$}{\it{Department of Physics and Astronomy, University of
      Pennsylvania, Philadelphia, PA }}}
\email{\tt jaguirre@sas.upenn.edu}

\affil{{$^\casa$}{\it{CASA, University of Colorado, CB 389, Boulder, CO 80309}}}
%, \email{John.Bally@casa.colorado.edu}
%      \email{adam.ginsburg@colorado.edu}
%      \email{cara.battersby@colorado.edu}
%      \email{Jason.Glenn@colorado.edu} \email{pmh@astro.as.utexas.edu}
%      \email{Guy.Stringfellow@colorado.edu} }}

\affil{{$^\utexas$}{\it{Department of Astronomy, University of Texas,
      1 University Station C1400, Austin, TX 78712}}}
%,            
%      \email{nje@astro.utexas.edu},
%      \email{nordhaus@astro.utexas.edu}}}

\affil{{$^\virginia$}{\it{Department of Astronomy, University of
Virginia, P.O. Box 400325, Charlottesville, VA 22904}}}
%, 
%    \email{meredith.drosback@origins.colorado.edu}}}

\affil{{$^\ucf$}{\it{Department of Physics, University of Central
Florida}}}
%,           
%                                              } ,   \email{tbradley@physics.ucf.edu}} }

\affil{{$^\wisc$}{\it{Department of Astronomy, University of Wisconsin,
       Madison, WI 53706}}}
% , \email{cyganow@astro.wisc.edu}}}

\affil{{$^\jpl$}{\it{Jet Propulsion Laboratory, California Institute
   of Technology, 4800 Oak Grove Dr., Pasadena, CA 91104}}}
%,
%       \email{cdd@submm.caltech.edu}}}

\affil{{$^{\ubc}$}{\it{Department of Physics and Astronomy, University
       of British Columbia, Okanagan }}}
% ,               \email{erik.rosolowsky@ubc.ca}}}

%\affil{{$^{\arizona}$}{\it{ Steward Observatory, University of
%Arizona, 933 North Cherry Avenue, Tucson, AZ 85721 }}}
%%\email{yshirley@as.arizona.edu} \email{wschlingman@as.arizona.edu}}}

\affil{{$^{\hawaiihilo}$}{\it{Institute for Astronomy, University of
        Hawaii,640 N. Aohoku Pl., Hilo, HI 96720}}}
%\email{joshw@ifa.hawaii.edu} }}

\affil{{$^{\hawaiimain}$}{\it{Institute for Astronomy, University of
Hawaii,680 Woodlawn Drive, Honolulu, HI 96822}}}
%\email{jpw@ifa.hawaii.edu}}

%\altaffiltext{\penn}{University of Pennsylvania, Philadelphia, PA 19104}
%%Jansky Fellow, National Radio Astronomy Observatory} 
%\altaffiltext{\casa}{CASA, University of Colorado, CB 389, Boulder, CO 80309}
%\altaffiltext{\utexas}{Department of Astronomy, University of Texas, 
%       1 University Station C1400, Austin, TX 78712}
%\altaffiltext{4}{Caltech Submillimeter Observatory, Hilo, HI}
%\altaffiltext{5}{Institute for Astronomy (IfA), University of Hawaii
%640 N. Aohoku Pl., Hilo, HI 96720}
%
\begin{abstract}

@@

@@

The advent of focal plane arrays containing hundreds of individual
bolometers sensitive to millimeter and sub-millimeter (sub-mm)
radiation has enabled large-scale surveys of mm-wavelength continuum
emission from Galactic interstellar dust.  The great advantage of
continuum observations at 1.1mm is that the mass column density does
not depend strongly on the dust temperature $T_d$ (varying only
linearly unless $T_d$ is less than about 10 K) and is independent of
the gas volume density.  Using the same dust opacity as
\citet{enoch06} and \citet{young06}, and assuming a gas-to-dust mass
ratio $X$ of 100, the mass sensitivity can be written as
\begin{equation} 
\label{eq:Mass}
M_{gas}\approx
0.22\left( e^{12.9/T_d}-1 \right) \left({S_\nu\over 15\; {\rm mJy}} \right)
D^{2}_{kpc} \msol 
\end{equation}
where $D_{kpc}$ is the distance to the source in kpc.  (At $S_\nu =
15$~mJy this corresponds to an RMS extinction sensitivity of $A_V\sim
1$ mag, with the same $T_d$ dependence).  It is thus possible with a
wide area, relatively shallow 1.1~mm survey to provide an unmatched,
uniform inventory of massive star-forming and starless condensations.

This survey presents the results of one of the first wide-area,
untargeted surveys of the Galactic Plane in the millimeter continuum.
We have used the 144 element Bolocam focal plane array, mounted at the
Cassegrain focus of the 10 meter diameter sub-mm telescope at the
Caltech Sub-millimeter Observatory (CSO), to survey over \bgpsarea\
square degrees of the northern Galactic plane at a wavelength of 1.1
mm in the dust continuum.  Figure \ref{fig:Coverage} shows the
coverage of the \bgps.  Maps of the entire BGPS are presented in
Figure \ref{fig:BGPSMontage}.  We have detected \ncores\ cores in the
surveyed fields, providing an unbiased sample of cores suitable for
multi-wavelength and high-resolution studies with existing telescopes
and future facilities.  The core catalog is described in a companion
paper \citep{rosolowsky09}.

@@

%\Figure{mass_limits}{The limiting gas mass of cores detectable
%with at a flux limit of 15 mJy RMS. Detection thresholds at 3 (dashed)
%and 5 (solid) $\sigma$ are shown for a range of dust temperatures.
%Dust temperatures $<20$~K are most interesting for this survey, since
%these are the sources most likely to be pre- or proto-stellar
%condensations.}{fig:mass_limits}{1.0}
%
\Figure{coverage_moll} {The coverage of the BGPS.  The background
greyscale is IRAS 100 \mum.  
%The Bolocam coverage is shown with white
%solid lines in the top two panels and black in the bottom three.
%Major regions of Galactic star formation are labeled.  The coverage of
%{\em Spitzer}-GLIMPSE \citep{benjamin03} and the FCRAO $^{13}$CO GRS
%\citep{jackson06} are shown.
}{fig:Coverage}{1.0}

% TEST
% Useful, but not here
%\begin{figure}
%  \subfigure[$l=-10.5$ to $l=19.5$ Units are Jy/Beam.  The brightest
%	sources, e.g. Sgr B2, Sgr A, and sources near $l=10$ and
%	$l=13$, appear to be saturated, but this is only a display
%	artifact.  The noise is more pronounced from $l=-7$ to $l=-2$
%	because this region was observed less.]{\label{fig:BGPSMontage-a}\includegraphics[scale=0.75]{fig1_pg1_bw_inv_bar_label}}\\
%\pagebreak
%\subfigure[$l=19.5$ to $l=49.5$.  Units are Jy/Beam .  G34.3+0.15, W
%	51, W 43, W 49, and M 17 appear to be saturated, but this is
%	only a display artifact.]
%{\label{fig:BGPSMontage-b}\includegraphics[scale=0.75]{fig1_pg2_bw_inv_bar_label}}
%
%\end{figure}
%% --------------------------------
%

\newcounter{subfig}
\setcounter{subfig}{1}
\renewcommand{\thefigure}{\arabic{figure}\alph{subfig}}

%\begin{sidewaysfigure}[h]
\begin{figure}
  \begin{minipage}{6.5in}
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg1_bw_inv_bar_label}
      \caption{$l=-10.5$ to $l=19.5$ Units are Jy/Beam.  The brightest
	sources, e.g. Sgr B2, Sgr A, and sources near $l=10$ and
	$l=13$, appear to be saturated, but this is only a display
	artifact.  The noise is more pronounced from $l=-7$ to $l=-2$
	because this region was observed less.}
    \end{center}
    \label{fig:BGPSMontage}
  \end{minipage}
\end{figure}

\addtocounter{figure}{-1}
\addtocounter{subfig}{1}

\begin{figure}
  \begin{minipage}{6.5in} 
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg2_bw_inv_bar_label} 
      \caption{$l=19.5$ to $l=49.5$.  Units are Jy/Beam .  G34.3+0.15, W
	51, W 43, W 49, and M 17 appear to be saturated, but this is
	only a display artifact}
    \end{center}
  \end{minipage}
\end{figure}

\addtocounter{figure}{-1}
\addtocounter{subfig}{1}

\begin{figure}
  \begin{minipage}{6.5in} 
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg3_bw_inv_bar_label}
      \caption{$l=49.5$ to $l=74.5$.}
    \end{center}
  \end{minipage}
\end{figure}

\addtocounter{figure}{-1}
\addtocounter{subfig}{1}

\begin{figure}
  \hspace{-1in}
  \includegraphics[scale=0.8]{fig1_pg5_bw_inv_bar_label} 
  \caption{The Cygnus Arm.}
\end{figure}
%  \vspace{-4in}

\addtocounter{figure}{-1}
\addtocounter{subfig}{1}

\begin{figure}
\hspace{-0.5in}
  \includegraphics[scale=0.8]{fig1_pg4_bw_inv_bar_label} 
  \caption{Cloud complexes centered at $l=111$ in the Perseus Arm.}
\end{figure}

\addtocounter{figure}{-1}
\addtocounter{subfig}{1}

\begin{figure}
  \hspace{-1in}
  \includegraphics[scale=0.8]{fig1_pg6_bw_inv_bar_label} 
  \caption{W3/4/5}
\end{figure}

\renewcommand{\thefigure}{\arabic{figure}}

@@

Bolocam\footnote{{\tt http://www.cso.caltech.edu/bolocam}} is the
facility 144-element bolometer array camera operating from the 10.4 m
dish of the Caltech Submillimeter Observatory (CSO) on the summit of
Mauna Kea.  We used the filter configuration with a bandcenter of
268~GHz (hereafter 1.1~mm) and fractional bandwidth $\Delta \nu/\nu =
0.17$.  The passband is constructed to exclude the $^{12}\mathrm{CO}(2
\to 1)$ emission line.  Figure \ref{fig:Bandpass} shows the passband.
We compute color corrections for sources of varying spectral index in
Appendix \ref{app:ColorCorrections}.

@@

\Figure{bolocam_bandpass}{The Bolocam 1.1 mm bandpass.  Bright
molecular emission lines are shown at their rest frequencies.  Note
that the Bolocam passband rejects $>90\%$ of the $^{12}$CO flux,
leaving our continuum measurements largely uncontaminated.  SO and
CH3OH lines are probably the dominant contributors to line flux in the
passband.  \citet{nummelin1998} found that 22\% of the flux in one
pointing towards Sgr B2 was from lines, and \citet{yoshida2005}
reported $>40\%$ in Orion A from line emission} {fig:Bandpass}{1.0}

@@

\FigureTwo{bolosens050904_o32}{scan_10deg}{Left: the position,
ellipticity and relative response of the detectors in the Bolocam
focal plane as mapped to the sky.  Right: The effect of field rotation
on the coverage obtained via Bolocam raster scans.}{fig:Array}{1.0}

\section{Flux Density and Surface Brightness Calibration}

The absolute flux calibration is obtained from observations of Mars
and Uranus
% no good observations, per AG
%and Neptune 
(the ``primary calibrators'').  The millimeter-wave flux of these
planets is known to $\sim 5\%$ \citep{orton86,griffin93}.
Calibrations at (sub)millimeter wavelengths are strongly affected by
atmospheric opacity corrections.  Further, the detector responsivity
of Bolocam's bolometers is a non-linear function of the mean
atmospheric loading.

@@

A fit to the observed values for the primary calibrators was performed
for the each epoch separately.
%CHECK
The agreement between epochs was good, so a single combined
calibration was used for all data here.
%We have found that the calibration can differ
%between successive cooldowns of the instrument by $\sim10\%$ \Check\
%but is generally stable over the period of order a month which
%constitutes each of the observing epochs.  
The calibration curve is shown in Figure \ref{fig:CalibrationCurves}.  The
resulting error on the calibration curve fit is less than 
%CHECK
$5\%$ (statistical) over the observed range of $\tau$.  A tally of all
contributions to the flux density error is given in Table
\ref{tab:ErrorBudget}.

The above flux density calibration only accounts for the average
calibration of bolometer Volts to Janskys.  It is also necessary to
account for the variation of bolometer response across the focal
plane.  We do this by monitoring the response to the atmosphere
emission in all bolometers.  Since the atmosphere is common to all
detectors, any changes in response are due to the intrinsic properties
of the detectors.  The change in relative response with loading for a
typical detector is shown in Figure \ref{fig:CalibrationCurves}.  The
actual application of the the method to the time series of the
bolometer data is shown in Figure \ref{fig:Flatfield}, and is
discussed further in Section \ref{sec:Mapping}.

The instrument beam size (PSF) is measured using the planets Uranus
and Neptune which are nearly point sources for Bolocam.
%The semi-diameter of
%Mars over this period ranged from 2.3\arcsec\ to 2.9\arcsec, with
%Neptune $\sim1.1\arcsec$.  Relative to the \bcamfwhm~FWHM Bolocam
%beam, these may be regarded as point sources; the finite size of Mars
%results in a $< 1.5\%$ correction of the observed flux.
The maps of the planets, including all detectors, are fit to a
Gaussian profile.  The beam area is that averaged over all detectors,
and is ?$1.71\times10^{-6}$\TBD\ ($35.6$\farcs) steradians.  This
allows conversion of the maps, calibrated in Jy beam$^{-1}$, into
surface brightness (MJy sr$^{-1}$).  The uncertainty in the beam size
is ?$.7\%$\TBD\ (number quoted is the uncertainty in the mean beam
size), leading to an uncertainty in the surface brightness calibration
of \TBD.

@@

Comparison of the accuracy of the Bolocam flux density calibration
with the SCUBA 850 \mum\ results is complicated by the unknown effect
of the dust spectral index.  However, there are several large-area
surveys with the MAMBO instrument at 1.2 mm with which we can compare.
Figure \ref{fig:CygnusComparison} shows the comparison of fluxes
extracted from our map in the Cygnus Arm with those of
\citet{motte07}.

\FigureTwo{fluxcal_errorbars}{rel_flux_cal_050303_ob3_to_050405_o66}
%{calibration_plot_2004}
{Left: Average calibration curve [V/Jy] versus the mean detector
voltage, a proxy for atmospheric loading.  Red points are observations
of Mars and black of Uranus.  Right: Scaling of the
relative calibration with atmospheric loading for bolocam
143. {\em To be replaced with new figure}}{fig:CalibrationCurves}{1.0}

\Figure{relsens_cal}{An illustration of the relative sensitivity
calibration using the atmosphere as a calibrator.  Black is the median
over all bolometers (the atmosphere model), red and green are
individual bolometers. a. before and b. after applying the relative
calibration.  Note the improved agreement.}{fig:Flatfield}{1.0}

\Figure{cygnus_flux_comparison}{A comparison of the fluxes extracted
in the same aperture from the \citet{motte07} MAMBO map and the
Bolocam map of this paper.}{fig:CygnusComparison}{1.0}

@@

We constructed an absolute reference system for the BGPS by observing
bright quasars and blazars near the Galactic Plane.  The distribution
of sources over the sky is shown in Figure
\ref{fig:PointingCalibrators} and the list is given in Appendix
\ref{app:PointingCalibrators}.

@@

For the purposes of comparison with other surveys, we present in
Appendix \ref{app:ReferenceSources} a selection

@@

\FigureTwo{plane_astrometry}{pointing_model_skydist}{Left: The
distribution of absolute reference sources (crosses) in the Galactic
Plane (Appendix \ref{app:PointingCalibrators}).  Also shown (crosses)
are the locations of bright, isolated, compact sources for comparison
with other surveys, as given in Appendix \ref{app:ReferenceSources}.
Right: The distribution of the pointing calibration sources across the
local sky in Hawaii.  Notice the good sampling of the entire region
used for the Survey.}{fig:PointingCalibrators}{1.0}

\begin{figure}
  \subfigure[The pointing model correction (left) and residuals
(right) for Epoch V, from which the master reference images were
derived for subsequent alignment. {\em To be replaced with newer
figure?}]
{\label{fig:PointingModel-a}\includegraphics[scale=1.0]{pointing_model}}\\

  \subfigure[The residuals of the pointing model.  Note the Gaussian
distribution.]{\label{fig:PointingModel-b}\includegraphics[scale=0.7]{pointing_model_residuals}}

\caption{Pointing model.}
\end{figure}

@@

We proceed iteratively as follows. 
\begin{equation}
\label{eq:FirstIterStep}
\tilde{a}^{(n)}(t) = 
\sum_i{d_i(t)-\tilde{s}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)}
\end{equation}
where $\tilde{s}^{(0)}_i(t)=0$.  This mean atmosphere model is then
fit to each bolometer to obtain the relative gains (flat field) of the
detectors
\begin{equation}
\tilde{a}^{(n)}_i(t) = \frac{d_i(t) \cdot a(t)}{d_i(t)^2} a(t) = r_i a(t)
\end{equation}
This is illustrated in Figure \ref{fig:Flatfield}.  The instrument
errors, which are typically smaller than the atmosphere, are then
estimated
\begin{equation}
\tilde{e}^{(n)}_i(t) = some models
\end{equation}
and both the atmosphere and instrument models are subtracted from the
raw time series
\begin{equation}
d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)
\approx s_i(t) + \epsi_i(t)
\end{equation}
This is the best estimate of signal plus irreducible noise at
iteration step $n$, and is made into a map
\begin{equation}
M[d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)] = \tilde{m}^{(n)}
\end{equation}
The current best map $\tilde{m}^{(n)}$ is then deconvolved to provide
a relatively low-noise, smooth map from which to generate a timestream
\begin{equation}
T[{\cal D}[\tilde{m}^{(n)}]] = \tilde{s}^{(n)}_i(t)
\end{equation}
We note that it is possible to produce an ``error map''
\begin{equation}
N = M[d_i(t) - \tilde{s}^{(n)}_i(t) - \tilde{a}^{(n)}_i(t) - 
\tilde{e}^{(n)}_i(t)]
\end{equation}
The process then begins again with Equation \ref{eq:FirstIterStep}.

@@

\Figure{flagger_withglitch}{An illustration of the flagging process
using the waterfall plot. The glitch apparent at middle right is the
same one whose time series is shown in Figure
\ref{fig:CosmicRayGlitch}.  Because of the PCA subtraction, the effect
of the glitch also propagates over the other bolometers, appearing as
a horizontal stripe.}  {fig:Flag1}{1.0}

@@

\section{Table of Absolute Pointing Sources}
\label{app:PointingCalibrators}
